---
title: "IODS final assignment"
author: "Ilari Hotti"
output:
  html_document:
    theme: united
    toc: true
    toc_depth: 2
    fig_caption: true
    fig_width: 8
    fig_height: 6
    code_folding: hide
---


***
## Abstract


***
## Hypothesis


***
## Presenting the data

#### Basic info
I chose a survey dataset from [kaggle](https://www.kaggle.com/miroslavsabo/young-people-survey) for my final assignment.

The data is from a Slovakian survey about music and movie preferences, hobbies & interests, phobias, health habits, personality traits, views on life , opinions, spending habits and finally demographics. Participants were Slovakians aged 15 to 30. Most of the variables had integer values from 1 to 5 representing how much the participant agrees with the statement in question.


#### Data wrangling
The dataset origanally contained 1010 observations and 150 variables. I created 3 new variables: `BMI` (calculated from weight and height), `Sports` (averaged passive and active sports interests) and `Depression` (averaged 10 questions that indicated depressed and anxious thinking: Self.criticism, Hypochondria, Loneliness, Changing.the.past, Life.struggles, Getting.up, Happiness, Energylevels, Personality_positivity and Socializing (last five had to be flipped around)). I also created a categorical variable for participants who scored higher or lower than 3.3 in my `Depression` index. 

I chose to include all the variables about music and movie preferences, `Healthy_eating`, `Dancing`, `Age` and `Gender` in addition to the variables I created. So in the end there are 39 variables.

There were also missing values and removing them meant the dataset shrank a bit to 832 observations.

#### Examining data structure
```{r warning=FALSE, message=FALSE}
library("tidyverse")
library("ggplot2")
library("dplyr")
library("GGally")
library("corrplot")
library("gridExtra")
ym <- read.table("data/young_minds.txt")
Palette <- c("#56B4E9", "firebrick1")
Palette2 <- c("#56B4E9", "gray27")
Palette3 <- c("#56B4E9", "orange")
Palette4 <- c("#56B4E9", "purple")
sexpalette = c("blue", "deeppink")
```

Let's examine the data set. 

```{r warning=FALSE, message=FALSE, fig.height=4}
glimpse(ym)
summary(ym$Gender)
summary(ym$Depressed)
age <- ggplot(ym, aes(x = Age)) + geom_histogram(binwidth = 1, color = "black", fill = "#56B4E9")
sex <- ggplot(ym, aes(x = "" , y  = Gender, fill=Gender)) + geom_bar(width = 1, stat = "identity") + coord_polar("y", start=0) + scale_fill_manual(values=sexpalette)

grid.arrange(age, sex, ncol=2)


```

There are 497 female and 335 male participants. 129 of them are depressed according to my variable and categorization. Most are around 20 years old.

#### Visualizing the variables

```{r warning=FALSE, message=FALSE}
Mu <- ggplot(ym, aes(x = Music)) + geom_histogram(binwidth = 1, color = "black", fill = "#56B4E9")
Mo <- ggplot(ym, aes(x = Movies)) + geom_histogram(binwidth = 1, color = "black", fill = "#56B4E9")
Da <- ggplot(ym, aes(x = Dancing)) + geom_histogram(binwidth = 1, color = "black", fill = "#56B4E9")
Sp <- ggplot(ym, aes(x = Sports)) + geom_histogram(binwidth = 1, color = "black", fill = "#56B4E9")
Bm <- ggplot(ym, aes(x = BMI)) + geom_histogram(binwidth = 0.3, fill = "firebrick1") + xlim(15, 40)
De <- ggplot(ym, aes(x = Depression)) + geom_histogram(binwidth = 0.1, fill = "darkblue")
grid.arrange(Mu, Mo, Da, Sp, Bm, De, ncol=2)
``` 

Everybody likes music and movies in general. 


### Correlations

So what kind of correlations can we find in the data? We can have a short look at the correlations between different music preferences and movie preferences.



```{r warning=FALSE, message=FALSE}
music_matrix <- cor(ym[9:27]) %>% round(digits = 2)
movie_matrix <- cor(ym[28:39]) %>% round(digits = 2)

corrplot(music_matrix, method="circle", type="lower", cl.pos = "b", tl.pos = "d", tl.cex = 0.8)
corrplot(movie_matrix, method="circle", type="lower", cl.pos = "b", tl.pos = "d", tl.cex = 0.8)
```

The correlations make sense but are not very interesting. 
```{r warning=FALSE, message=FALSE}
spoD <- ggplot(ym, aes(x = Sports, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette)
eatD <- ggplot(ym, aes(x = Healthy_eating, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette)
danD <- ggplot(filter(ym, Gender == "female"), aes(x = Dancing, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette) +xlab("Dancing(girls)")
danDb <- ggplot(filter(ym, Gender == "male"), aes(x = Dancing, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette) +xlab("Dancing(boys)")
bmiD <- ggplot(ym, aes(x = BMI, fill = Depressed)) + geom_histogram(binwidth = 0.5 , position="fill") + scale_fill_manual(values=Palette)+xlim(15,40)
ageD <- ggplot(ym, aes(x = Age, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette)
grid.arrange(spoD, eatD, danD, danDb, bmiD, ageD, ncol=2)
```

```{r warning=FALSE, message=FALSE, fig.height=3, fig.width=5}
genD <- ggplot(ym, aes(x = Gender, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette)
genD
```


How about music and movie preference correlations with `Depression`?

```{r warning=FALSE, message=FALSE}
cor(ym[9:27], ym[4]) %>% round(digits = 2)
cor(ym[28:39], ym[4]) %>% round(digits = 2)
```

Depressed people seem to prefer slow, classical and rock music, and dislike rap and dance. As a side note the boys' `Musical` variable has opposite correlation with `Depression` to girls' one.

In the movie section highest positive correlation is with `Romantic` and highest negative correlation with `Action`. 

Some barcharts of the most interesting variables:

```{r warning=FALSE, message=FALSE, fig.height=3}
musicalboys <- ggplot(filter(ym, Gender == "male"), aes(x = Musical, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette4) + xlab("Musical(boys)") 
musicalgirls <- ggplot(filter(ym, Gender == "female"), aes(x = Musical, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette4) + xlab("Musical(girls)")
grid.arrange(musicalboys, musicalgirls, ncol=2)

```



```{r warning=FALSE, message=FALSE}
slowD <- ggplot(ym, aes(x = Slow_vs_Fast, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette2)
danceD <- ggplot(ym, aes(x = Dance, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette2)
classicalD <- ggplot(ym, aes(x = Classical, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette2)
rapD <- ggplot(ym, aes(x = Hiphop_Rap, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette2)
rockD <- ggplot(ym, aes(x = Rock, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette2)
altD <- ggplot(ym, aes(x = Alternative, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette2)
grid.arrange(slowD, danceD, classicalD, rapD, rockD, altD, ncol=2)
```
```{r warning=FALSE, message=FALSE, fig.height=2}
romD <- ggplot(ym, aes(x = Romantic, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette3)
actD <- ggplot(ym, aes(x = Action, fill = Depressed)) + geom_bar(position="fill") + scale_fill_manual(values=Palette3)
grid.arrange(romD, actD, ncol=2)
```



Could there be statistical significance to any explanatory connection in the data regarding the `Depression` variable?

*** 
## Fitting a regression model
```{r warning=FALSE, message=FALSE}
m_music <- glm(Depressed ~ Slow_vs_Fast + Hiphop_Rap + Dance, data = ym, family = "binomial")
summary(m_music)
```

```{r warning=FALSE, message=FALSE}
m <- glm(Depressed ~ Sports + Healthy_eating + Gender, data = ym, family = "binomial")
summary(m)
```

So `Hiphop_Rap` is a statistically important explanatory variable.

```{r warning=FALSE, message=FALSE}
m_music2 <- glm(Depressed ~ Rock + Classical + Alternative, data = ym, family = "binomial")
summary(m_music2)
```

Intercept = 3* ???

```{r warning=FALSE, message=FALSE}
m_movies <- glm(Depressed ~ Romantic + Tales + Animated, data = ym, family = "binomial")
summary(m_movies)

m_movies2 <- glm(Depressed ~ Action + Western, data = ym, family = "binomial")
summary(m_movies2)

m_bonus <- glm(Depressed ~ Action + Hiphop_Rap + Sports, data = ym, family = "binomial")
summary(m_bonus)
m_bonus2 <- glm(Depressed ~ Action, data = ym, family = "binomial")
summary(m_bonus2)
```
***
##Analysis of the model



***
## Conclusions

